Greenhouse gas emissions model

### 2021-08-19 210617_GHG_model.Rmd
first running model (needs quality checks)

# the following is page 10.41 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories
mm_ch4<-(volsol*365)*
        (Bo*0.67*
           ((mms_pasture*MCFpasture)+
            (mms_spread*MCFspread)+
            (mms_drylot*MCFdrylot)+
            (mms_solidstore*MCFsolidstore)+
            (mms_composted*MCFcomposted)+
            (mms_liquid*MCFliquid)+
            (mms_biogas*MCFbiogas)+
            (mms_burn*MCFburn)+
              # the original code was lacking the 'sold' fraction - we added this for completeness
            (mms_sold*MCFsold)))

@@

##### #####
##### 
##### WE STOPPED HERE
#####
##### #####

# Calculate Nitrous oxide emissions from manure and urine on pasture
# ### ##### CW Note #### 
# ### ##### To get this to run I am assuming a temp 10% pregnancy_coefficient_Cp
pregnancy_coefficient_Cp <- 0.1
### #####
### ####
n_intake<-(gross_energy_intake/18.45)*((pregnancy_coefficient_Cp/100)/6.25)
n_retension<-((milk_yield*(3.3/100))/6.38)
n_excretion<-n_intake*(1-n_retension)*365
mm_direct_n2o<-n_excretion*((mms_pasture*0.02)+(mms_spread*0.005)+(mms_drylot*0.02)+(mms_solidstore*0.005)+
               (mms_composted*0.006)+(mms_liquid*0.005)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28)

# Nitrous oxide (N2O) from volatization 
mm_vol_n2o<-(n_excretion*((mms_pasture*0.02)+(mms_spread*0.07)+(mms_drylot*0.2)+(mms_solidstore*0.3)+
            (mms_composted*0.3)+(mms_liquid*0.4)+(mms_biogas*0)+(mms_burn*0))*0.01*(44/28))

# Nitrous oxide (N2O) from leaching 
mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)

# Estimate CO2eq for manure nitrous
mm_CO2_N2Od<-mm_direct_n2o*298*365
mm_CO2_N2Ovol<-mm_vol_n2o*298*365
mm_CO2_N2Oleach<-mm_leach_n2o*298*365
mm_CO2_N2O<-mm_CO2_N2Od+mm_CO2_N2Ovol+mm_CO2_N2Oleach

# Emissions from feed production and transport. 
feed_CO2<-feed_prod_CO2*feed_trans_CO2

@@

# Relationship between greenhouse gas intensity and milk yields. 
# Fat adjusts
ann_adj <-kg_FPCM_year*365 
df <- data.frame(
  GHGi = on_farm/ann_adj, 
  FPCM = ann_adj)

@@

return(list(enteric_CH4=enteric_CH4,
            #CW NOTE ####
            # ####working backwards from 'enteric_CO2' I see that 
            # #### this is the product of something twice multiplied by 365
            # ####### 
            enteric_CO2=enteric_CO2,
            mm_CO2=mm_CO2,
            mm_CO2_N2O=mm_CO2_N2O,
            feed_CO2=feed_CO2,
            on_farm=on_farm))
}

>>>>

# decisions on manure management
# the management options we're considering here are the ones from the Excel table. The IPCC guidelines
# list more and partly different ones.

# We're deciding on manure management according to a decision tree (Cory will build a figure):

# question 1: how much of the manure is centrally collected? share_centrally
# question 2: of manure that's not centrally collected, how much is burned? share_central_burn
mms_burn<-(1-share_centrally)*share_central_burn
mms_pasture<-(1-share_centrally)*(1-share_central_burn)

# question 3: of centrally collected manure, how much is spread daily on cropland or pasture? share_spread
mms_spread<-share_centrally*share_spread
not_spread<-share_centrally*(1-share_spread)
# question 4: how much of the manure that's not spread is sold? share_sold
mms_sold<-not_spread*share_sold
not_sold<-not_spread*(1-share_sold)
# question 5: how much of the stored manure that's not sold is stored in liquid form? share_liquid
liquid<-not_sold*share_liquid
solid<-not_sold*(1-share_liquid)
# question 6: how much of the liquid manure is converted to biogas? share_biogas
mms_biogas<-liquid*share_biogas
mms_liquid_slurry<-liquid*(1-share_biogas)
# question 7: how much of the solid manure is composted? share_compost
mms_composted<-solid*share_compost
remain_solid<-solid*(1-share_compost)
# question 8: how much of the solid manure is stored in dry lots? share_drylot
mms_drylot<-remain_solid*share_drylot
mms_solidstore<-remain_solid*(1-share_drylot)

# the following is page 10.41 of the 2006 IPCC Guidelines for National Greenhouse Gas Inventories
mm_ch4<-(volsol*365)*
        (Bo*0.67*
           ((mms_pasture*MCFpasture)+
            (mms_spread*MCFspread)+
            (mms_drylot*MCFdrylot)+
            (mms_solidstore*MCFsolidstore)+
            (mms_composted*MCFcomposted)+
            (mms_liquid_slurry*MCFliquid)+
            (mms_biogas*MCFbiogas)+
            (mms_burn*MCFburn)+
              # the original code was lacking the 'sold' fraction - we added this for completeness
            (mms_sold*MCFsold)))

@@

# Calculate Nitrous oxide emissions from manure and urine on pasture

### ####
n_intake<-(gross_energy_intake/18.45)*((perc_crude_protein_diet/100)/6.25) # equation 10.32

# There was some code here on N retention in milk. That's apparently from equation 10.33, but it only had the first part of the equation, missing the part on weight gain. We decided to go with the alternative way of estimating N retention (figures given in Table 10.20)
# n_retention<-((milk_yield*(3.3/100))/6.38)

# Table 10.20 contains N retention coefficients:
# Dairy cows: 0.20 N_retention_dairy_cow
# Other cattle: 0.07 N_retention_other_cattle

animal_type_dairy_cow=1 ###ASSUMPTION

if(animal_type==animal_type_dairy_cow)
  n_excretion<-n_intake*(1-N_retention_dairy_cow)*365 else
    n_excretion<-n_intake*(1-N_retention_other_cattle)*365 #equation 10.31

# Direct nitrous oxide emissions (Eq 10.25 from IPCC guidelines)
mm_direct_n2o<-n_excretion*((mms_pasture*mndec_pasture)+
                           (mms_spread*mndec_spread)+
                           (mms_drylot*mndec_drylot)+
                           (mms_solidstore*mndec_solidstore)+
                           (mms_composted*mndec_composted)+
                           (mms_liquid*mndec_liquid)+
                           (mms_biogas*mndec_biogas)+
                           (mms_burn*mndec_burn)+
                           (mms_sold*mndec_sold))*(44/28)

# Nitrous oxide (N2O) from volatization (Eq 10.26 from IPCC guidelines)
mm_vol_n2o<-n_excretion*((mms_pasture*mnvc_pasture)+
                           (mms_spread*mnvc_spread)+
                           (mms_drylot*mnvc_drylot)+
                           (mms_solidstore*mnvc_solidstore)+
                           (mms_composted*mnvc_composted)+
                           (mms_liquid*mnvc_liquid)+
                           (mms_biogas*mnvc_biogas)+
                           (mms_burn*mnvc_burn)+
                           (mms_sold*mnvc_sold))*(44/28)

# Nitrous oxide (N2O) from leaching
# comment: this is probably wrong, since leached nitrogen doesn't emit (much) nitrous oxide
#mm_leach_n2o<-n_excretion*((mms_pasture*0.3)+(mms_spread*0.3)+(mms_drylot*0.3)+(mms_solidstore*0.3)+
#              (mms_composted*0.3)+(mms_liquid*0.3)+(mms_biogas*0)+(mms_burn*0))*0.0075*(44/28)

# Estimate CO2eq for manure nitrous
mm_CO2_N2Od<-mm_direct_n2o*298*365 # (298 is the CO2-equivalent of N2O)
mm_CO2_N2Ovol<-mm_vol_n2o*298*365
# mm_CO2_N2Oleach<-mm_leach_n2o*298*365
mm_CO2_N2O<-mm_CO2_N2Od+mm_CO2_N2Ovol #+mm_CO2_N2Oleach

# Emissions from feed production and transport. 
# comment: the numbers in the table seem spuriously precise.
feed_CO2<-feed_prod_CO2*feed_trans_CO2

@@

# Relationship between greenhouse gas intensity and milk yields. 
# Fat adjusts
#ann_adj <-kg_FPCM_year*365 
#df <- data.frame(
#  GHGi = on_farm/ann_adj, 
#  FPCM = ann_adj)

@@

return(list(enteric_CH4=enteric_CH4,
            #CW NOTE ####
            # ####working backwards from 'enteric_CO2' I see that 
            # #### this is the product of something twice multiplied by 365
            # ####### 
            milk_yield=milk_yield,
            enteric_CO2=enteric_CO2,
            mm_CO2=mm_CO2,
            mm_CO2_N2O=mm_CO2_N2O,
            feed_CO2=feed_CO2,
            on_farm=on_farm))
}
